Reproducible Research -- Assignment 1
========================================================

# Loading and preprocessing the data

```{r, echo=TRUE}
setwd('C:/Users/Stephen/Desktop')
mydataFull <- read.csv('activity.csv')
mydata <- droplevels(mydataFull[!is.na(mydataFull$steps), ])
```

# Total number of steps per day

## Histogram

```{r fig.height=4.5, echo=TRUE}
stepsPerDay <- tapply(mydata$steps, mydata$date, sum)
hist(stepsPerDay)
```

## Mean and median of total steps per day

```{r, echo=TRUE}
cat('Mean of daily total steps: ', mean(stepsPerDay), '\n', sep = '')
cat('Median of daily total steps: ', median(stepsPerDay), '\n', sep = '')
```

# Daily activity pattern

```{r fig.height=4.5, echo=TRUE}
meanEachIntvl <- tapply(mydata$steps, as.factor(mydata$interval), mean)
plot(meanEachIntvl ~ unique(mydata$interval), type = 'l', main = 'Daily activity pattern')
```

# Time interval with maximum mean steps per day

```{r, echo=TRUE}
cat('Time interval with maximum mean steps: ', unique(mydata$interval)[which(meanEachIntvl==max(meanEachIntvl))], sep = '')
```

# Missing values

## Number of missing values

```{r, echo=TRUE}
cat('Number of missing values: ', sum(is.na(mydataFull$steps)), sep = '')
```

## Input missing values and re-analysis

```{r fig.height=4.5, echo=TRUE}
mydataMissing <- droplevels(mydataFull[is.na(mydataFull$steps), ])
mydataMissing$steps <- rep(meanEachIntvl, 8)
mydataFull <- rbind(mydataMissing, mydata)
stepsPerDayFull <- tapply(mydataFull$steps, mydataFull$date, sum)
hist(stepsPerDayFull)
cat('Mean of daily total steps (inputed data): ', mean(stepsPerDayFull), '\n', sep = '')
cat('Median of daily total steps (inputed data): ', median(stepsPerDayFull), '\n', sep = '')
```

## Impact of inputation

Because the missing data were inputed using mean steps within a certain 5-minute interval, the inputation strategy did not affect the **mean** of daily total steps; however, the **median** of daily total steps had a tiny increase.

# Weekdays vs Weekends

## Data preparation

```{r, echo=TRUE}
weekday <- weekdays(as.Date(mydataFull$date), abbreviate = T)
weekday[weekday == 'Sat' | weekday == 'Sun'] <- 'weekend'
weekday[weekday != 'weekend'] <- 'weekday'
mydataFull$weekday <- weekday
Weekdays <- droplevels(mydataFull[mydataFull$weekday == 'weekday', ])
Weekends <- droplevels(mydataFull[mydataFull$weekday == 'weekend', ])
meanEachIntvlWeekdays <- tapply(Weekdays$steps, as.factor(Weekdays$interval), mean)
meanEachIntvlWeekends <- tapply(Weekends$steps, as.factor(Weekends$interval), mean)
```

## Time series plots

```{r fig.height=4.5, echo=TRUE}
plot(meanEachIntvlWeekdays ~ unique(Weekdays$interval), type = 'l', main = 'Weekdays')
plot(meanEachIntvlWeekends ~ unique(Weekends$interval), type = 'l', main = 'Weekends')
```
